Concurrent Programming with Harmony

### 2020-08-28 book/paper.tex
Last

\texttt{release\_wlock} executes \texttt{notifyAll}(?\textit{wcond}) as well as
\texttt{notify}(?\textit{wcond)}.
Again, because we do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter, or, alternatively,
up to one waiting writer can enter.  So \texttt{release\_wlock} wakes up all
potential candidates.
There are two things to note here.  First, unlike split binary semaphores or Hoare
monitors, where multiple waiting readers would have to be signaled one at a time in a
baton-passing fashion (see \autoref{fig:RWsplitsema}), with Mesa monitors
all readers are awaked in one fell swoop using \texttt{notifyAll}.
Second, both readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \textbf{while} loop, re-checking the condition that they
are waiting for.  So if both type of processes are waiting, either all the readers
get to enter next or one of the writers gets to enter next.

Much of the complexity of programming with Mesa condition variables is in figuring out
when to invoke \texttt{notify} and when to invoke \texttt{notifyAll}.  As a rule of
thumb: be conservative---it is better to wake up too many processes than too few.
Waking up too many processes may lead to some inefficiency, but waking up to few
may cause the application to get stuck.  Harmony can be particularly helpful here, as
it examines every corner case.  You can safely try to replace each \texttt{notifyAll}
with \texttt{notify} and see if every possible execution of the the application still
terminates.

>>>>

\texttt{release\_wlock} executes \texttt{notifyAll}(?\textit{rcond}) as well as
\texttt{notify}(?\textit{wcond)}.
Again, because we do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter, or, alternatively,
up to one waiting writer can enter.  So \texttt{release\_wlock} wakes up all
potential candidates.
There are two things to note here.  First, unlike split binary semaphores or Hoare
monitors, where multiple waiting readers would have to be signaled one at a time in a
baton-passing fashion (see \autoref{fig:RWsplitsema}), with Mesa monitors
all readers are awaked in one fell swoop using \texttt{notifyAll}.
Second, both readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \textbf{while} loop, re-checking the condition that they
are waiting for.  So if both type of processes are waiting, either all the readers
get to enter next or one of the writers gets to enter next.

Much of the complexity of programming with Mesa condition variables is in figuring out
when to invoke \texttt{notify} and when to invoke \texttt{notifyAll}.  As a rule of
thumb: be conservative---it is better to wake up too many processes than too few.
Waking up too many processes may lead to some inefficiency, but waking up too few
may cause the application to get stuck.  Harmony can be particularly helpful here, as
it examines every corner case.  You can safely try to replace each \texttt{notifyAll}
with \texttt{notify} and see if every possible execution of the the application still
terminates.
